### 2012-10-12 paper/paper.tex
paper commit

%\documentclass[11pt,letterpaper,onecolumn]{article}
\documentclass{llncs}
\usepackage{llncsdoc}
%\usepackage[left=1in,top=1in,right=1in,bottom=1in,nohead]{geometry}

\usepackage{epsfig}
\usepackage{times}
\usepackage{latexsym}
\usepackage{graphicx}
\usepackage{ltxtable}
\usepackage{tabularx}
\usepackage{hhline}
\usepackage{color,colortbl}
\usepackage{url}
\usepackage{boxedminipage}
\usepackage{hyperref}
\usepackage{breakurl}
\usepackage{comment}
%\usepackage{fullpage}

\newcommand\T{\rule{0pt}{2.6ex}}
\newcommand\B{\rule[-1.2ex]{0pt}{0pt}}
\begin{document}

\title{Contextual Identity: A Framework for Usable Privacy}

%\author{Monica Chew,Sid Stamm\\
%\email{\{mmc,sid\}@mozilla.com}}
%\institute{Mozilla}
\maketitle
\section{Introduction}
\begin{quote}I am large, I contain multitudes. --- Walt Whitman,
\begin{em}Song of Myself\end{em} \end{quote}

People reveal different aspects of themselves depending on context. At any
given time, they may be acting as a friend, relative, spouse, co-worker,
acquaintance, or stranger. This observation is not new; Erving Goffman called
this impression management, and Carl Jung called it persona
theory~\cite{goffman,jung}. In this short paper, we call it contextual identity.
%identity as social construct?

Helen Nissenbaum has long argued that privacy violations come not from the
simple sharing of personal information, but rather from sharing that
information in a way that violates social norms, or in the wrong
context~\cite{nissenbaum}. The desire for spontaneous, positive human
interaction often necessitates sharing personal information. Take, for example,
PatientsLikeMe~\cite{patientslikeme}. This site functions as a support group
and also a place for people to share health information, such as drug regimens,
treatment history, side-effects, and more. Though this information is
traditionally considered extremely sensitive, the users of this site clearly
consider the benefits of sharing (longitudinal graphs of their own health
history, building community with people who share their disease) to outweigh
the privacy risks. So long as the contextual integrity of the shared data is
kept intact, users will continue to trust and use the site. Selling user data
intended for one purpose for another (e.g., to insurance companies to discover
uninsurable patients) would be a clear violation of that expectation.
%this story is a positive story to shock security folk with the breadth and
%depth of personal information that ordinary users will share
% instead we could lead with a shocking, negative story like how gay youth get
% killed when they're outed

In offline environments, managing contextual identities is more intuitive than
in online environments. There are cues to help you determine where you are,
who your intended audience is, how many people will overhear you, and how
likely your information is to be re-broadcast in a different context. It is
also true that in offline environments, humans don't have perfect memories and
will eventually what forget.  However, with the advent of increasingly vast
searching, indexing, and archiving cabilities, one cannot rely on forgetting in
online enviroments~\cite{delete}. Thus, managing contextual identities in
online environments has become increasingly difficult and fraught with
mistakes.

One possible solution to this problem is never to post personal information
online, or at least to consider the entire world might be the audience.
However, this solution defies human nature. Given that there are 1 billion
active Facebook users and that 30-40\% of spoken communication is devoted to
informing others about ourselves, humans are bound to fall short of these
guidelines~\cite{tamir,fbusers}. The security and privacy community should
recognize this and develop techniques to meet users' needs rather than
expecting human nature to change.

We attempt to formalize the notion of contextual identity, use it as a
framework to understand existing privacy violations, and propose mitigations to
those violations.

\section{Definitions}
Let
\begin{itemize}
\item $R$ be a set of actors and their relationships to the user
\item $T$ be a set of tasks and their side-effects when performed in service of
$R$
\item $P$ an impression that actors in $R$ have of the user when performing tasks in $T$
\end{itemize}
Let a \textbf{contextual identity} $I$ be a loose coupling of $(R, T, P)$.
A user may have multiple of these. These identities change over time, as users
form and dissolve relationships, and focus on different sets of tasks. The
desirable outcome is that the user can accomplish tasks in $T$ while being able
to predict $R$ and $P$.
%predict rather than "control" is deliberate, since a user can't realistically
%expect to engage in behavior that violates social norms and give whatever
%impression they want

This definition focusses on relationships because most privacy concerns stem
from causing problems with people the users knows~\cite{wang}. This definition
focusses on tasks because users are task-oriented, and it is easy to map user
actions (page navigation, sending email, posting, filling forms) to tasks.
(include?) The definitions of $R$ and $P$ are deliberately underspecified, as
this is work in progress.

\section{Previous work}

Many authors have discovered attacks to link different contextual identities to
the same person. Narayanan and Shmatikov present a re-identification technique
to merge anonymized social graphs from different networks and prove that
Netflix and IMDB and Flickr and Twitter can be
linked~\cite{narayanan1,narayanan2}. Lindamood et al.~and Mislove et al.~show
how to infer previously undisclosed information from released social networking
data~\cite{lindamood,mislove}.

Many authors have noted how re-broadcasting information information out of
context and making information discoverable can lead to user distress (even if
the information was previously public, but not easily
discoverable)~\cite{boyd1,chew,nissenbaum}.

Barth et al.~ provide a formal model for contextual integrity of user
data~\cite{barth}. In contrast to this work, the contextual integrity model
focusses on data as principal rather than users.

\begin{comment}TODO: figure out how to cite this. Why is this short
paper's model worth pursuing? It's simpler, and focusses on the user rather
than data.
\end{comment}

Although network attacks are very powerful, at the current time they require a
level of expertise and resources not typically held by those causing the most
worry to users. A brief survey of social networking advice reveals that the
major causes of concern come not from a dedicated graph expert, but people that
the user knows: friends, family, and co-workers~\cite{fbtips2,fbtips1}.
Surveys conducted by Wang et al.~show that typical regrets from posting on
Facebook stem almost exclusively from fear of negative interactions with people
that the users knows~\cite{wang}. The consequences of these negative
interactions can lead to loss of employment or breaking personal relationships.
By contrast, much privacy research has focussed on network privacy, where the
bad actor is a behavioral tracking service, or a state agent.

\section{Contextual identity violations}
\label{sec:examples}
\subsection{Redistributing information out of context}
Re-distributing information out of its original context often leads to
embarrassment~\cite{nissenbaum}.  In November 2007, Facebook Beacon allowed
third-party sites to publish purchases, travel bookings, movie rentals to the
user's activity stream.  Because of poor opt-out and lack of visibility into
what was being published, user outcry was immediate~\cite{mccarthy,nakashima}.
In December 2007, Google Reader exposed RSS feeds of user-marked news stories
to the user's Google Talk contacts\footnote{Google Talk contacts are everyone
with whom a user has chatted, which might include co-workers, supervisors, and
friends}. Although this feed was always public, prior to this launch it was not
discoverable, leading to a bad experience for many users~\cite{helft}. In
September 2012, Facebook imported old wall posts from 2008 into the new
Timeline interface.  Although wall posts were always visible from profile
pages, the new Timeline interface brought old wall posts (which users used to
treat as private messages, before the advent of ``Like'' and comment buttons)
to the attention of an audience that the user never intended~\cite{timeline}.

In all these examples, the audience $R$ of the identity expanded without user
intervention.

\subsection{Unsatisfiable policies}
Sometimes service providers have policies which preclude isolating multiple
identites. For example, Facebook and Google have a ``Real Names'' policy, which
requires users to register for accounts with their legal
name~\cite{fb_names,google_names}.  These policies presume the users can be one
person to everyone. These policies ignore that community-building happens in
many different contexts, that users have many legitimate reasons for presenting
different identities in different contexts, and that users don't necessarily
want those identities to be linked. For example, disallowing avatar handles as
a primary identifier makes building a gaming community difficult at best, and
linking a gaming identity to a professional identity has already caused
problems for one exposed has already caused problems for Maine Senate candidate
and World of Warcraft gamer Colleen Lachowicz in 2012~\cite{maine}.

It is impossible for users who want to isolate multiple contextual identities
to participate in these networks without violating the terms of service.

\subsection{Federated login}
Facebook Connect is a login platform that allows third-party websites to
authenticate users using their Facebook identity~\cite{fb_connect}. Because it
is against the terms of service to have multiple Facebook accounts, using
Facebook Connect has the unwanted side-effect of linking multiple contextual
identities.

Other login platforms and protocols, including OAuth and BrowserID
do not suffer from this policy or design error~\cite{browserid,oauth}.
\begin{comment}
(TODO:
this is actually kinda complicated -- yahoo, twitter, google allow multiple
accounts but the multi-login support is broken at google, not sure how it is
with others, oauth protocol supports multiple accounts, persona clearly
supports multiple accounts)
\end{comment}

\subsection{User interface errors}
Even in the absence of changes made by service providers to re-contextualize
personal information, it is all too easy for users to broadcast information to
an unintended audience. This phenomenon is so common on Twitter that it has its
own name, ``DM fail'', or Direct Message fail: when the user posts a public
message instead of a private, direct message. Representative Anthony Weiner was
a victim of this mistake when he inadvertently sent compromising pictures of
himself publicly~\cite{weiner}. Considering that this mistake requires
mistyping a single character (\texttt{@} instead of \texttt{d}), it's no
surprise that DM failures are exceedingly common.

Similar to DM failures, posting to the wrong account is also a common mistake.
Because many jobs require posting on social networks on behalf of the company,
many people now have multiple accounts for personal and business use.
KitchenAid, Chrysler, and Google are just three examples whose employees have
made this mistake in the past year~\cite{kitchenaid,chrysler,yegge}.

\subsection{Third party tracking}
\label{sec:tracking}
Even if a user does not participate in social networks, they are still subject
to tracking by virtue of browsing the web. Third-party cookies, flash cookies,
web bugs, device fingerprinting, geolocation, history sniffing and more are
techniques that a dedicated third party could use to collect browsing history
and link contextual identities.

\section{Possible mitigations}

\subsection{Inferring contextual identities}
Accurately detecting the user's current contextual identity, whether
heuristically or through asking the user, is paramount to helping the user
manage it.

Although a person's contextual identities may be subconsciously clear,
expecting users to enumerate their identities and manually curate them is
probably too burdensome. Firefox has a Profile Manager, but not many people
know about it or how to use it, to the extent that it will be
removed\footnote{See \url{https://bugzilla.mozilla.org/show\_bug.cgi?id=214675\#c53}}.

Tang et al.~show that most people do not take the time to create labels and
maintain friend lists~\cite{tang}. However, the same authors show how to use
heterogenous networks to infer relationship types (e.g., manager-subordinate)
through the graph itself is unlabeled.

It may also be possible to infer tasks from the user's activity. For example,
the \texttt{about:profile} Firefox extension uses the Open Directory Project
and Alexa to categorize URLs that the user has visited~\cite{aboutprofile}.

\subsection{Automatic mitigations for information leakage}
Given the set of relationships $R$ and tasks $T$, we can categorize a user's
contextual identities and prevent information leakage between them.  For
example, preventing third-party cookies from being read and set across task
boundaries means that targeted ads for dating would not follow the user when
visiting a technology news site. Knowing the user's current contextual identity
also allows the browser to suppress URL bar suggestions that aren't relevant to
that identity, protecting the user from both shoulder-surfing and distraction.

Privacy modes in Firefox, Chrome, Safari, and Opera are a good first step, but
are no means a complete solution~\cite{ABBJ10}. There is no standard behavior
for private browsing mode, and thus interaction with extensions, treatment of
cookies, history, and bookmarks upon entering and exiting this mode are
different and serve different use cases. Depending on user desire, $R$ when
using private browsing mode may be other household members, or it may be a
service to which the user wants to limit cookie leakage. Another deficit is
that private browsing mode does not adequately serve long-term tasks (such as
curating a set of adult entertainment sites) in any major browser.

Another good partial solution is to prompt the user when they're about to link
multiple contextual identities through a federated login service. For example,
Persona is the suite of tools that implements the BrowserID protocol, and it
supports multiple email identifiers~\cite{browserid}. In this case, when the
user decides to register for a service using Persona, the task $T$ they are
performing is authentication, the relationships $R$ are ones they will engage
in when using the service, and $P$ is tied to the email identifier they use to
register for the service. Because we can categorize the type of service by URL
as \texttt{aboutprofile} does, Persona could prompt the user when they're about
to link the same email identifier to radically different contextual identities,
such as ones for dating and professional use.

\begin{comment}
TODO: Discuss how private browsing mode maps to R, T, P?

At home use case:
R is household
T is porn?
P is null

Searching use case:
R is Google
T is search
P is null

GoogleShring use case
R is Google and the GoogleSharing proxy
T is search or browsing (analytics) (can't use any services that require logon)
P is random
\end{comment}

\subsection{Recovering from mistakes}
\subsubsection{Auditing}
A major cause of user embarrassment is exposing information to the wrong
audience, as illustrated in section~\ref{sec:examples}. A browser is in the
perfect position to intermediate social network posts, aggregate them locally,
and present them to the user when they desire to audit their digital footprint.
For example, a user could see all of the comments and posts they made in the
last week and redact content that is sensitive or too negative, topics that are
frequently regretted~\cite{wang}. Although this approach does not help when the
service provider decides to change the audience on the user's behalf, auditing
gives the user a good opportunity to review posts that might cause
regrets. In the case of high-volume users, such a tool could even incorporate
sentiment analysis to aid the user in finding particularly problematic content.

Profiles can be generated by user-supplied application-level data, but also
implicitly through browsing via techniques in~\ref{sec:tracking}. Collusion is a
tool for visualizing cookies~\cite{collusion}, in particular the third-party
cookies set by tracking sites which are typically ad networks.  Once the user
knows which sites are tracking them, they can choose to block cookies from
those sites.

\subsubsection{Remembering to forget}
Humans don't have perfect memories, and neither should social
networks~\cite{delete}. Many users already engage in manual auditing and
deletion of old posts~\cite{fbtips2}. A better solution would be to build tools
that let the user manage this more easily. Both Twitter and Facebook provide
APIs for deletion\footnote{At the time of this writing, Google Plus offers a
read-only API}. One way to reduce the likelihood of linked contextual
identities would be to provide applications that expire posts after some time.

\begin{comment}
\cite{viegas}blogger's expectation of privacy
\end{comment}

\subsection{Preventing mistakes}
\subsubsection{Slowing information flow}
Sometimes a user might want to prevent mistakes from being made in the first
place. If a user is concerned about a particular service provider, they can
install an extension like Disconnect to disallow cookies from that provider
unless they explicitly want to interact with that site~\cite{disconnect}.

Protocols like Do Not Track are also a good first step to solving this
problem~\cite{dnt}. Although Do Not Track is unenforceable from the client
side, it provides a clear signal to service providers and ad networks
that the user does not want to be tracked. As adoption increases, ignoring this
signal will lead to loss of reputation for service providers that explicitly
act against their users' wishes.

\subsubsection{Visual reminders of context}
To mitigate privacy violations caused by poor user interfaces, the onus must be
on software developers to provide sufficient visual cues of context. For
example, many but not all Google products support multiple accounts. However,
the visual cue to indicate which account is active is a small, easy to ignore
text name at the top right of the page.

Much work has been done for visual verification of domain authenticity in the
context of phishing, but not much has been done for visual verification of
which user credentials are active~\cite{skins}.

\begin{comment}
first party regrets, third party regrets
impossible to judge how big friends of friends
\end{comment}

\section{Open problems}
The mitigations proposed in this work become more complex in the face of shared
devices and overlapping identities. There are good reasons to want to intermix
work and personal identities (increased exposure, serendipitous connections)
and so many people use the same account for both personal and work reasons.

Mobile growth is driving most internet growth, and much of that growth comes
developing countries. The literature on how device sharing works in those
markets is completely lacking. Managing contextual identities on shared devices
promises to be a difficult problem.

\section{Discussion}
People have multiple contextual identities. Sharing personal information in
those contexts is a great way to foster postive human interaction.  We hope
that contextual identity will serve as a framework for both developers to
understand their users' privacy needs, and to develop tools that allow users to
act in their contextual identities in a spontaneous, relaxed, and interesting
way.

\begin{comment}
papers:
expectation of privacy\cite{viegas}
http://jcmc.indiana.edu/vol10/issue3/viegas.html

danah boyd
making sense of privacy and publicity: whose voice counts? does your child's
teacher have a right to be public in her other roles online? does she have a
right to be secular? religious? friend? lover? if you work for company X and
you disagree with their policies, are you allowed to be public about
that:?\cite{boyd2}

close to you “stranger on a train” “stranger on a plane”
\end{comment}

\begin{comment}
\section{Acknowledgements}
The authors thank
Ben Adida
Ed Lee
Lucas Adamski
Mike Connor

\end{comment}

\bibliographystyle{plain}
\bibliography{paper}
\end{document}